Unveiling Robustness Threats by Mining the Exception Stack Traces of over 600 Android Projects

### 2014-09-05 android-stacks.tex
results - intro e wrappings

\begin{abstract}
The number of mobile apps is increasing in a daily rate. If on the one hand they
extend phones capabilities far beyond of the basic calls, on the other hand they have to 
cope with an increasing number of exceptinal conditions 
(e.g., faults in underlying middleware or hardware; memory and battery restrictions). 
Therefore, mechanisms for exception detection and handling are not an optional add-on but a 
fundamental part of such apps. However, studies have shown that such mechanisms are the less 
understood and tested parts of the system. As a consequence they may affect the system the 
other way around, making the system even less robust, leading to failures such as uncaught exceptions 
and unintended handling (i.e., exceptions being mistakenly caught by existing handlers).
This paper reports a study which mined the stack traces embedded on 
159,048 issues reported on Android projects hosted in GitHub (482 projs.) 
and Google Code (157 projs.). 
The goal of this study was to: investigate the whether the exception stack traces
 can reveal common characteristics on the exception-related code that
may be favoring the introduction of the aforementioned failures, which would be responsible for 
decreasing the application robustness. We call such characteristics ``fragilities" of the exception-related code.
Overall 6,005 exception stack traces were extracted, and analyzed in 
combination with source code and bytecode analysis. 
Some fragilities consistently detected in this study include: 
(1) undocumented runtime exceptions thrown by third party code;
(2) undocumented checked exceptions thrown by JNI native code;
(3) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions),
revealing that Java hybrid exception model is not fully used according to its purpose.
Moreover, about 50\% of all analyzed projects reported at least one NullPointerExeption
as the root cause of an exception stack trace.
This study provide important insights on how such fragilities impose challenges 
to application robustness not only in Android environment but also to other 
systems based on Java exception model.

@@

@@

Along the recent years, we have witnessed an astonishing increase in the number 
mobile applications. Such increase is motivated by pervasiveness of the World Wide Web 
and the unprecedented increase of global smartphone shipments -
the mobile industry shipped over 315 million smartphones in the last quarter of 2013 ~\cite{googleio}.

In this context, Android has become a leading platform - thanks to its open-source 
model and hardware partners like Samsung, HTC, Motorola, Asus (besides Nexus Google's
 own device) ~\cite{gartner}.  Google estimates that there is over 1 billion Android active users per month ~\cite{googleio},
and the number of apps built on top of Android increases in a daily rate.

If on the one hand such applications extend phones capabilities 
far beyond of the basic calls and textual messages, on the other hand
they have to face an increasing number of threats to application robustness
 arising from: faults in underlying middleware and hardware, 
faults on third party services and libraries; compatibility issues; 
faults on connected devices (e.g. camera, GPS); memory and 
battery restrictions; among others.

Therefore, techniques for error detection and handling are not  an optional add-on but a 
fundamental part of such apps.
% as any other modern system~\cite{bruntink2006discovering}. 
The Android platform uses the exception handling mechanism embedded
 in Java to deal with such conditions. Although the exception handling mechanism
~\cite{goodenough1975exception} is one of the most used schemes for
detecting and recovering from such exceptional conditions, studies (e.g.~\cite{miller1997issues,Robil00,shah2010understanding, garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}),
have shown that it is generally poorly understood and the less tested parts of the system ~\cite{coelho2011unveiling}.
As a consequence they may affect the system the other way around.
The exception-related code may favor the introduction of failures such as uncaught exceptions ~\cite{jo2004uncaught, zhang2012amplifying}
and unintended handling ~/cite{miller1997issues} (i.e., exceptions being mistakenly caught by existing handlers) 
which make the system even less robust.

Our conjecture is that several Android applications may be suffering from common 
fragilities on the exception-related code (of the applications themselves or 
the underlaying platform) which may be degratind their robustness.
We call ``fragilities" characteristics on the exception-related code that favor the introduction
of aforementioned failures.

@@

Until recent years, the availability of data, reporting such failures, was scarce. 
It comprised only few large scale projects (e.g. the Mozilla crash report dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
Recently we have observed a rapid increase in the number of open
source projects that make their version control systems as well as the issue tracking systems
publicly available in repository hosting sites such as GitHub and Google Code.

When available, the exception stack traces provide a useful source of information about system crashes ~\cite{bettenburg2008makes} which 
 can enable different kinds of post-mortem analysis and support:  debugging~\cite{schroter2010stack}, 
 bug classification and clustering ~\cite{wang2013improving, kim2011crash, dhaliwal2011classifying},  
automated bug fixing ~\cite{sinha2009fault} and fault-proneness prediction models ~\cite{kim2013predicting}.

@@

\begin{table*}
  \centering
  \begin{tabular}{rcccccccc}
    \hline
    \bfseries{Root Exception} &  \multicolumn{2}{c}{\bfseries{Projects (GH\&CG)}} &  \multicolumn{2}{c}{\bfseries{Occurrences (GH\&CG)}} & \textsf{Android} & \textsf{Libcore} & \textsf{App} & \textsf{Lib} \\
    & \bfseries{\#} &  \bfseries{\%} & \bfseries{\# } & \bfseries{\% } &&&&\\
    \hline

java.lang.NullPointerException	&	332	&	51,96\%	&	1664	&	27,71\%	&	525	&	20	&	836	&	280	\\
java.lang.IllegalStateException	&	120	&	18,78\%	&	278	&	4,63\%	&	185	&	31	&	41	&	39	\\
java.lang.IllegalArgumentException	&	142	&	22,22\%	&	353	&	5,88\%	&	195	&	12	&	95	&	44	\\
java.lang.RuntimeException	&	122	&	19,09\%	&	319	&	5,31\%	&	203	&	2	&	64	&	51	\\
java.lang.OutOfMemoryError	&	78	&	12,21\%	&	237	&	3,95\%	&	141	&	16	&	35	&	34	\\
java.lang.NoClassDefFoundError	&	67	&	10,49\%	&	94	&	1,57\%	&	10	&	0	&	46	&	37	\\
java.lang.ClassCastException	&	64	&	10,02\%	&	130	&	2,16\%	&	55	&	0	&	55	&	20	\\
java.lang.IndexOutOfBoundsException	&	62	&	9,70\%	&	166	&	2,76\%	&	53	&	0	&	93	&	18	\\
java.lang.NoSuchMethodError	&	54	&	8,45\%	&	80	&	1,33\%	&	10	&	0	&	56	&	14	\\
java.util.ConcurrentModificationException	&	43	&	6,73\%	&	65	&	1,08\%	&	5	&	0	&	46	&	13	\\
\\

    \hline
  \end{tabular}
\caption{Root Exceptions occurrences and popularity in repositories hosted in Google Code (GC) and GitHub(GH).}
\label{tab:topten}
\end{table*}

@@

We can observe that most of the exceptions on this list are implicitly thrown by the
 runtime environment due to programming mistakes  (e.g., out-of-bounds array index, division-by-zero, access to a null reference)
 or resource limitation (e.g., OutOfMemoryError). 
From this set the java.lang.NullPointerException was the most reported root cause (27,71\% ). 
If we consider the frequencly of NullPointerException 
across projects, we can observe that 51,96\% of all projects reported at least one exception stack 
trace on which the NullPoiterException was the root cause.
 The NullPointerExceptions are mainly signaled inside the applicaiton code and the Android platform,
 although we also find NullPointerExceptions being signaled from third-party libraries. Regarding
reusable code, there is no consensus whether it is a good or bad practice to 
re-throw a NullPointerException. Some prefer to encapsulate such an exception on
IllegalArgumentException, while others~\cite{bloch2008effective} argue that the
NullPointerException makes the cause of the problem explicit and hence 
should not be wrapped.

@@

Table ~\ref{tab:causes} summarized the results. It also illustrates the number of general Java exceptions - 
whose full name does not carry the reason for the exception to be thrown.
To ensure the quality of the analysis, three independent coders classified a randomly selected
sample of 25 exception types (from the total 100) using the same list of concerns;
the interater agreement was high (96\%). We could also observe that
We can therefore assume that the categories are descriptive 
enough for covering the range of exception types analyzed.

@@

\begin{table}
  \centering
  \begin{tabular}{lr}
    \hline
    \bfseries{Cause} &  \bfseries{\% Occurrences} \\
    \hline
      Programming Mistakes &  52,0\%\\ 
      %Resources Restriction   &   23,9\% \\ 
      General Exceptions (Error, Exception, Runtime)    &  6,7\%\\
    \hline
 \end{tabular}
  \caption{Identifying the concerns related to root exceptions}
  \label{tab:causes}
\end{table}

@@

Table~\ref{tab:wrappingandroid} presents all wrappings found in this study, that
envolve different exception types (i.e., Error, Checked, Exception). We call 
such a wrapping a ``cross-type wrapping".  Such wrappings reveals 
violations to best practices I and II (e.g. Error wrapping a Checked exception),
To illustrate the problem behind such wrappings, we can use one of the wrappings found in this study. 
When the developer is confronted with a checked exception, the designer of the API is telling him 
to handle the exceptional condition, however we could observe that in some cases the 
checked exception wrapped an OutOfMemoryError, which represent resource deficiency detected 
by the runtime environment which make impossible the program to recover. 
Hence, trying to handle this exception may lead the program to an unpredictable state.

Next we discuss and exemplify the cross-type wrappings
 found in this study.

@@

@@

\emph{\textbf{Checked wrapping Runtime Exception.}} Manually inspecting the stack 
traces we could observe that most of the wrappings were performed 
by RoboSpice (an android library that supports the creation of asynchronous long running 
tasks, which has 1.682 stars and 367 forks in GitHub) and by the reflection Java library (used by Android, 
Application and library classes). Inspecting the code related to such wrappings we could observe that
the reflection library convert every exception that happens 
when invoking a reflected method into an InvoncationTargetException (this conversion is 
performed inside native code).

@@

\emph{\textbf{Multiple Wrappings.}} We could also observe that in 205 stacks found
 in Android repository the exception was wrapped  more then once. One interesting 
multiple-wrapping found in Android repo was the 
following: Runtime-Checked-Runtime-Checked-Runtime-Checked-Runtime.

AS WE COULD OBSERVE...

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
MUDARRRR Odd wrappings were detected (e.g., a checked exception wrapping an OutOfMemoryError).
The study also detected an undocumented checked exception signaled by a native method defined in Android.
}
\end{minipage}}

@@

\begin{figure*} \centering \includegraphics[scale=0.55]{codeexamples.pdf}
\caption{Code snippets illustrating fragilities on the exception-related code} \label{fig:snippets} \end{figure*}

@@

\subsection{The exception handling confusion problem}
When (mis)applied, the exception wrapping can make the exception-related code
 more complex and lead to what we call an \emph{exception handling confusion problem}.
Such problem can lead the program to an unpredictable state in the presence of exceptions. 
To illustrate this problem, we can use one of the examples found in this study: when the developer is 
confronted with a checked exception, the designer of the API is telling him 
to handle the exceptional condition, however we could observe that in some cases the 
checked exception wrapped an OutOfMemoryError, which represent resource deficiency detected 
by the runtime environment which make impossible the program to recover.

Currently there is no way of enforcing Java exception type conventions during program development.
Hence, further investigation is needed on finding ways to help developers in dealing with such
 problem, either preventing odd wrappings or enabling the developer to better deal with them or even
researches on the real usefulness of Java hybrid exception model.

@@

\section{Threats to Validity}
\label{sec:threats}
\noindent\emph{Internal Validity} We used a heuristics-based parser to mine
exceptions from issues.  Our parsing strategy was conservative by default; for
example, we only considered exception names using a fully qualified class name
as valid exception identifiers, while, in many cases, developers use the
exception name in issue description. Conservative parsing may minimize false
positives, which was our initial target, but also tends to increase false
negatives, which means that some cases may have not been identified as
exceptions or stack traces. Our limited manual inspection did not reveal such
cases.

\noindent\emph{External Validity} The results presented here were based on mining
issues on Github through the GHTorrent dataset. While comprehensive and
extensive, GHTorrent is not an exact replica of Github, so several issues might
be left out. Due to the way data collection works with GHTorrent, for projects
that are relatively inactive, GHTorrent might have not collected a significant
proportion of their issues. We did not investigate the extend of this threat on
our sample.
\footnote{to do: OLHAR ARTIGO DO FSE QUE TRABALHA COM OPENSOURCE.
FALAR QUE DESCOBERTAS FORAM DE LIBS JAVA. APESAR DO ESTUDO NAO SER FEITO EM PROJS JAVA.....
PODEMOS CONCLUIR ALGUMAS COISAS QUE SE APLICAM A TODOS JAVA}

\noindent\emph{Construct Validity} Parts of our analysis are based on the availability of stack traces on issues on
Github. In using this dataset, we make an underlying assumption: the
stack traces reported on issues are representative of real crashes in
the applications. The assumption is impossible to mitigate without access to
the full set of crash data per application. Some services exist to collect 
crash data from mobile applications ~\cite{BugSe14,BugSn14,Googl14,Acra14}.
However, such services do not provide open access to the crash reports
of their client applications. Having access to such a dataset would allow 
a replication of our analysis on an even wider setting.

\section{Related Work}
\label{sec:rele}

@@

http://docs.oracle.com/javase/7/docs/api/

%JAVA.LANG EXCEPTIONS
ArithmeticException
ArrayIndexOutOfBoundsException
ArrayStoreException
ClassCastException
ClassNotFoundException
CloneNotSupportedException
EnumConstantNotPresentException
Exception
IllegalAccessException
IllegalArgumentException
IllegalMonitorStateException
IllegalStateException
IllegalThreadStateException
IndexOutOfBoundsException
InstantiationException
InterruptedException
NegativeArraySizeException
NoSuchFieldException
NoSuchMethodException
NullPointerException
NumberFormatException
ReflectiveOperationException
RuntimeException
SecurityException
StringIndexOutOfBoundsException
TypeNotPresentException
UnsupportedOperationException

JAVA UTIL:
ConcurrentModificationException
DuplicateFormatFlagsException
EmptyStackException
FormatFlagsConversionMismatchException
FormatterClosedException
IllegalFormatCodePointException
IllegalFormatConversionException
IllegalFormatException
IllegalFormatFlagsException
IllegalFormatPrecisionException
IllegalFormatWidthException
IllformedLocaleException
InputMismatchException
InvalidPropertiesFormatException
MissingFormatArgumentException
MissingFormatWidthException
MissingResourceException
NoSuchElementException
TooManyListenersException
UnknownFormatConversionException
UnknownFormatFlagsException

>>>>

\begin{abstract}
The number of mobile apps is increasing in a daily rate. If on the one hand they
extend phones capabilities far beyond of the basic calls, on the other hand they have to 
cope with an increasing number of exceptinal conditions 
(e.g., faults in underlying middleware or hardware; memory and battery restrictions). 
Therefore, mechanisms for exception detection and handling are not an optional add-on but a 
fundamental part of such apps. However, studies have shown that such mechanisms are the less 
understood and tested parts of the system. As a consequence they may affect the system the 
other way around, making the system even less robust, leading to failures such as uncaught exceptions 
and unintended handling (i.e., exceptions being mistakenly caught by existing handlers).
This paper reports a study which mined the stack traces embedded on 
159,048 issues reported on Android projects hosted in GitHub (482 projs.) 
and Google Code (157 projs.). 
The goal of this study was to: investigate the whether the exception stack traces
 can reveal common characteristics on the exception-related code that
may be favoring the introduction of the aforementioned failures, which would be responsible for 
decreasing the application robustness. We call such characteristics ``fragilities" of the exception-related code.
Overall 6,005 exception stack traces were extracted, and analyzed in 
combination with source code and bytecode analysis. 
Some fragilities consistently detected in this study include: 
(1) undocumented runtime exceptions thrown by third party code;
(2) undocumented checked exceptions thrown by JNI interface;
(3) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions),
revealing that Java hybrid exception model is not fully used according to its purpose.
Moreover, about 50\% of all analyzed projects reported at least one NullPointerExeption
as the root cause of an exception stack trace.
This study provide important insights on how such fragilities impose challenges 
to application robustness not only in Android environment but also to other 
systems based on Java exception model.

@@

%Such increase is motivated by pervasiveness of the World Wide Web 
%and the unprecedented increase of global smartphone shipments -
%the mobile industry shipped over 315 million smartphones in the last quarter of 2013 ~\cite{googleio}.

%In this context, Android has become a leading platform - thanks to its open-source 
%model and hardware partners like Samsung, HTC, Motorola, Asus (besides Nexus Google's
% own device) ~\cite{gartner}.  Google estimates that there is over 1 billion Android active users per month ~\cite{googleio},
%and the number of apps built on top of Android increases in a daily rate.

%Our conjecture is that several Android applications may be suffering from common 
%fragilities on the exception-related code (of the applications themselves or 
%the underlaying platform) which may be degratind their robustness.
%We call ``fragilities" characteristics on the exception-related code that favor the introduction
%of aforementioned failures.

%In Java, when an application fails due to an uncaught exception, 
%it automatically terminates, while the system prints a stack trace to the console, 
%or log file ~\cite{gosling2000java}.  A typical Java stack trace consists of  the fully qualified name 
%the thrown exception and the ordered list of methods that were active on the call stack before 
%such exception has occurred ~\cite{gosling2000java,bloch2008effective}.
%and unintended handling~\cite{miller1997issues} (i.e., exceptions being mistakenly caught by existing handlers)

@@

Along the recent years, we have witnessed an astonishing increase in the number 
mobile applications. If on the one hand such applications extend phones capabilities 
far beyond of the basic calls and textual messages, on the other hand
they have to face an increasing number of threats to application robustness
 arising from: faults on underlying middleware and hardware (e.g., GPS receiver, camera, sensors);
faults on third party services and libraries; compatibility issues ~\cite{McDon13}; 
 memory and battery restrictions~\cite{Zhang12}; concurrency issues ~\cite{ama2012}; among others.

Therefore, techniques for error detection and handling are not  an optional add-on but a 
fundamental part of such apps. The exception handling mechanism ~\cite{goodenough1975exception},
embedded in many mainstream programming languages, such as Java, C++ and C\#,
 is one of the most used techniques for detecting and recovering from such exceptional conditions. 
Android, the leadling plaform for mobile applications~\cite{gartner} (thanks to its open-source 
model and hardware partners like Samsung, HTC, Motorola, Asus besides Nexus Google's
 own device)  reuses the embedded Java Exception model to detect and handle 
 exceptional conditions.

Studies have shown, however, that the exception-related code is generally poorly understood and the
 less tested parts of the system~\cite{miller1997issues,Robil00,shah2010understanding, 
garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}.
As a consequence they may affect the system the other way around.
The exception-related code may favor the introduction of failures such as 
uncaught exceptions~\cite{jo2004uncaught, zhang2012amplifying}
which make the system even less robust~\cite{coelho2011unveiling}.

@@

Until recent years, the availability of data, reporting such failures, was scarce. 
It comprised only few large scale projects (e.g. the Mozilla crash report dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
Recently we have observed a rapid increase in the number of open
source projects that make their version control systems as well as the issue tracking systems
publicly available in repository hosting sites such as GitHub and Google Code.
Developers and users of open-source projects hosted these sites 
constantly uses the provided infra-strucute to share information across 
the software development (including exception stack traces detected).

%When available, the exception stack traces provide a useful source of information about system crashes ~\cite{bettenburg2008makes} which 
% can enable different kinds of post-mortem analysis and support:  debugging~\cite{schroter2010stack}, 
% bug classification and clustering~\cite{wang2013improving, kim2011crash, dhaliwal2011classifying},  
%automated bug fixing~\cite{sinha2009fault} and fault-proneness prediction models~\cite{kim2013predicting}.

@@

\begin{table*}
  \centering
  \begin{tabular}{rcccccccc}
    \hline
    \bfseries{Root Exception} &  \multicolumn{2}{c}{\bfseries{Projects}} &  \multicolumn{2}{c}{\bfseries{Occurrences}} & \textsf{Android} & \textsf{Libcore} & \textsf{App} & \textsf{Lib} \\
    & \bfseries{\#} &  \bfseries{\%} & \bfseries{\# } & \bfseries{\% } &&&&\\
    \hline

java.lang.NullPointerException	&	332	&	51,96\%	&	1664	&	27,71\%	&	525	&	20	&	836	&	280	\\
java.lang.IllegalStateException	&	120	&	18,78\%	&	278	&	4,63\%	&	185	&	31	&	41	&	39	\\
java.lang.IllegalArgumentException	&	142	&	22,22\%	&	353	&	5,88\%	&	195	&	12	&	95	&	44	\\
java.lang.RuntimeException	&	122	&	19,09\%	&	319	&	5,31\%	&	203	&	2	&	64	&	51	\\
java.lang.OutOfMemoryError	&	78	&	12,21\%	&	237	&	3,95\%	&	141	&	16	&	35	&	34	\\
java.lang.NoClassDefFoundError	&	67	&	10,49\%	&	94	&	1,57\%	&	10	&	0	&	46	&	37	\\
java.lang.ClassCastException	&	64	&	10,02\%	&	130	&	2,16\%	&	55	&	0	&	55	&	20	\\
java.lang.IndexOutOfBoundsException	&	62	&	9,70\%	&	166	&	2,76\%	&	53	&	0	&	93	&	18	\\
java.lang.NoSuchMethodError	&	54	&	8,45\%	&	80	&	1,33\%	&	10	&	0	&	56	&	14	\\
java.util.ConcurrentModificationException	&	43	&	6,73\%	&	65	&	1,08\%	&	5	&	0	&	46	&	13	\\
    \hline
  \end{tabular}
\caption{Root Exceptions occurrences and popularity in repositories hosted in Google Code (GC) and GitHub(GH).}
\label{tab:topten}
\end{table*}

@@

We can observe that most of the exceptions on this list are implicitly thrown by the
 runtime environment due to programming mistakes  (e.g., out-of-bounds array index, division-by-zero, access to a null reference)
 or resource limitation (e.g., OutOfMemoryError). 
From this set the java.lang.NullPointerException was the most reported root cause (27,71\% ). 
If we consider the frequencly of NullPointerException 
across projects, we can observe that 51,96\% of all projects reported at least one exception stack 
trace on which the NullPoiterException was the root cause.

 The NullPointerExceptions are mainly signaled inside the applicaiton code and the Android platform,
 although we could also find NullPointerExceptions being signaled by third-party libraries. 
Regarding reusable code, there is no consensus whether it is a good or bad practice to 
re-throw a NullPointerException. Some prefer to encapsulate such an exception on
IllegalArgumentException, while others~\cite{bloch2008effective} argue that the
NullPointerException makes the cause of the problem explicit and hence 
should not be wrapped.

@@

Table ~\ref{tab:causes} summarized the results. It also illustrates the number of general Java exceptions - 
whose full name does not carry the reason for the exception to be thrown.
To ensure the quality of the analysis, three independent coders classified a randomly selected
sample of 25 exception types (from the total 100) using the same list of concerns;
the resulting interrater agreement was high (96\%). We could also observe that
We can therefore assume that the exception names and Javadoc 
were enough for classifying the exception types analyzed.

@@

\begin{table}
  \centering
  \begin{tabular}{lr}
    \hline
    \bfseries{Cause} &  \bfseries{\% Occurrences} \\
    \hline
      Programming Mistakes &  52,0\%\\ 
      Memory Restriction   &   4,1\% \\ 
      General Exceptions   &  6,7\%\\
    \hline
 \end{tabular}
  \caption{Identifying the concerns related to root exceptions}
  \label{tab:causes}
\end{table}

@@

Table~\ref{tab:wrappingandroid} presents all wrappings found in this study, that
envolve different exception types (i.e., Error, Checked, Exception). We call 
such a wrapping a ``cross-type wrapping".  Some of these wrappings may reveal 
violations to best practices I and II (e.g. Error wrapping a Checked exception) as discussed next.

@@

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
An unexpected fragility was detected: a checked exception being thown and not specified on 
on the ``exception interface" of the signaler methods.
}
\end{minipage}}

@@

\emph{\textbf{Checked wrapping Runtime Exception.}} Manually inspecting the stack 
traces we could observe that such wrappings were performed either 
by RoboSpice (an android library that supports the creation of asynchronous long running 
tasks, which has 1.682 stars and 367 forks in GitHub) or by the reflection Java library (used by
methods defined in Android, Application and library classes).

@@

We could also observe that some stack traces included several intermediate cross-type wrappings.
One interesting example was the following: Runtime-Checked-Runtime-Checked-Runtime-Checked-Runtime.
Although some of these wrappings may be a result of design decision
the mis-use of exception wrapppings may make the exception handling 
code more complex (e.g., the multiple wrappings) and error-prone,
 and lead to what we call ``exception confusion problem"
To illustrate this problem we can use one of the wrappings discussed above.
When the developer is confronted with a checked exception, the designer of the API is telling him 
to handle the exceptional condition. However we could observe that in some cases the 
checked exception wrapped an OutOfMemoryError, which represent resource deficiency detected 
by the runtime environment which make impossible the program to recover. 
Hence, trying to handle this exception may lead the program to an unpredictable state.

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
Some of the detected cross-type exception wrappings points to a potential ``exception confusion problem".
To illustrate: a checked exception wrapping an OutOfMemoryError; and the multiple wrapping Runtime-Checked-Runtime-Checked-Runtime-Checked-Runtime.
}
\end{minipage}}

@@

\begin{figure*} \centering \includegraphics[scale=0.55]{examples_new2.png}
\caption{Code snippets illustrating fragilities on the exception-related code} \label{fig:snippets} \end{figure*}

@@

\subsection{The exception handling confusion problem}
When (mis)applied, the exception wrapping can make the exception-related code
 more complex and lead to what we call an \emph{exception handling confusion problem}.
Such problem can lead the program to an unpredictable state in the presence of exceptions,
as illustrated by the scenario on which a checked exception wrapped an OutOfMemoryError. 
Currently there is no way of enforcing Java exception type conventions during program development.
Hence, further investigation is needed on finding ways to help developers in dealing with such
 problem, either preventing odd wrappings or enabling the developer to better deal with them or even
researches on the real usefulness of Java hybrid exception model.

@@

\section{Threats to Validity}
\label{sec:threats}

\noindent\emph{Internal Validity} We used a heuristics-based parser to mine
exceptions from issues.  Our parsing strategy was conservative by default; for
example, we only considered exception names using a fully qualified class name
as valid exception identifiers, while, in many cases, developers use the
exception name in issue description. Conservative parsing may minimize false
positives, which was our initial target, but also tends to increase false
negatives, which means that some cases may have not been identified as
exceptions or stack traces. Our limited manual inspection did not reveal such
cases. Moreover, in this study we manually classified the exceptions 
related to programming mistakes. To ensure the quality of the analysis, 
we calculated the interrater agreement after three independent 
coders classified a randomly selected sample (of 25 exception 
types from the total 100); the interater agreement was high (96\%).

Another threat relates to the fact that parts of our analysis 
are based on the availability of stack traces on issues on Github and Googlecode. 
In using these datasets, we make an underlying assumption: the stack traces reported on issues are 
representative of valid crash information of the applications. 
One way to mitigate such threat would be to access to the full 
set of crash data per application. Although some services exist 
to collect crash data from mobile applications ~\cite{BugSe14,BugSn14,Googl14,Acra14},
they do not provide open access to the crash reports of their client applications.
In our study, we mitigated this threat by manually inspecting
the source code associated to a subset of reported exception stack traces.
This subset contained the stack traces related to the main findings 
of the study (e.g., `ùndocumented runtime and checked exceptions",
and cross-type wrappings"). Through the manual inspection
 we could map the stack traces information to real characteristics 
of the exception-related code of analyzed apps.

\noindent\emph{External Validity} First, we limited out analysis to open-source
Android apps. We are aware that the reported exceptions for commercial
apps can be different from the ones found in this study. However, 
in this study revealed that several reported exception stack traces 
were reporting root exceptions signaled by Android utility code and libraries. 
Moreover, since the ``fragilities" observed in this study are due to
characteristics of Java exception model, they impose challenges 
the robustness not ony to Android apps but also to other systems
 based on Java exception model.

Second, the results presented here were based on mining
issues on Github through the GHTorrent dataset. While comprehensive and
extensive, GHTorrent is not an exact replica of Github, so several issues might
be left out. Due to the way data collection works with GHTorrent, for projects
that are relatively inactive, GHTorrent might have not collected a significant
proportion of their issues. We did not investigate the extend of this threat on
our sample.

\footnote{to do: OLHAR ARTIGO DO FSE QUE TRABALHA COM OPENSOURCE.
FALAR QUE DESCOBERTAS FORAM DE LIBS JAVA. APESAR DO ESTUDO NAO SER FEITO EM PROJS JAVA.....
PODEMOS CONCLUIR ALGUMAS COISAS QUE SE APLICAM A TODOS JAVA}

\
\section{Related Work}
\label{sec:rele}

@@
